---
title: "Team Styles"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Purpose
The purpose of this notebook is to create 4 visualizations to share on the Atlantis platform. The 4 visualizations are as follows

1) Team Scoring (good, bad, fun, boring)

# Get Games for Analysis
The 2019-2020 Spanish La Liga season (ID:775) will be used to initially create each visualization. The 2019-2020 German Bundesliga will then be used to verify the functionalized process resulting from the initial creation. All games scheduled on or after 2020-06-10 will be removed. Only completed games will be used.

```{r}
source('src/sports/soccer/get_league_games.R')
id1 <- 775
rawGames1 <- get_league_games(leagueId = id1)
g1 <- rawGames1 %>%
	mutate(GameDate = lubridate::ymd(stringr::str_sub(string = GameTime, start = 1, end = 10))) %>%
	filter(GameDate < '2020-06-10') %>%
	select(HomeTeam, AwayTeam, HomeScore, AwayScore, HomeTeamLogoUrl, AwayTeamLogoUrl)
logos1 <- rawGames1 %>% select(Team = HomeTeam, Logo = HomeTeamLogoUrl) %>% distinct()
activeTeams1 <- c('Valencia' , 'Levante')
id2 <- 754
rawGames2 <- get_league_games(leagueId = id2)
g2 <- rawGames2 %>%
	mutate(GameDate = lubridate::ymd(stringr::str_sub(string = GameTime, start = 1, end = 10))) %>%
	filter(GameDate < '2020-06-10') %>%
	select(HomeTeam, AwayTeam, HomeScore, AwayScore, HomeTeamLogoUrl, AwayTeamLogoUrl)
logos2 <- rawGames2 %>% select(Team = HomeTeam, Logo = HomeTeamLogoUrl) %>% distinct()
activeTeams2 <- c('Union Berlin', 'Hertha Berlin')
```

# Team Scoring
The team scoring viz will use an X-Axis of ascending points scored and a Y-Axis of descending points scored. This will result in 4 quadrants:
- Good Team: Good Offense, Good Defense
- Bad Team: Bad Offense, Bad Defense
- Exciting Team: Good Offense, Bad Defense
- Boring Team: Bad Offense, Good Defense

This will be based on the average goals for/against for each team. Additionally, the highlighting of teams provided in a vector should be supported. 

First, the goal averages must be created. 

```{r 'Create Team Goal Averages'}
homeGoals <- g1 %>%
	group_by(HomeTeam) %>%
	summarise(GoalsFor = sum(HomeScore), GoalsAgainst = sum(AwayScore), Games = n()) %>%
	ungroup() %>%
	select(Team = HomeTeam, GoalsFor, GoalsAgainst, Games)
awayGoals <- g1 %>%
	group_by(AwayTeam) %>%
	summarise(GoalsFor = sum(AwayScore), GoalsAgainst = sum(HomeScore), Games = n()) %>%
	ungroup() %>%
	select(Team = AwayTeam, GoalsFor, GoalsAgainst, Games)
goals <- homeGoals %>%
	inner_join(awayGoals, by = 'Team') %>%
	mutate(GoalsFor = GoalsFor.x + GoalsFor.y,
		   GoalsAgainst = GoalsAgainst.x + GoalsAgainst.y,
		   Games = Games.x + Games.y,
		   AvgGoalsFor = GoalsFor / Games,
		   AvgGoalsAgainst = GoalsAgainst / Games) %>%
	inner_join(logos1, by = 'Team') %>%
	select(Team, AvgGoalsFor, AvgGoalsAgainst, Games, GoalsFor, GoalsAgainst, Logo) %>%
	arrange(-AvgGoalsFor, -AvgGoalsAgainst)
avgGoals <- mean(goals$AvgGoalsFor)
print(goals)
```

Next, create the base graph using ggplot
```{r}
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_point()
```

```{r 'Invert y-axis and add lines for 1:1 ratio and avg goals'}
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_point() +
	geom_abline(intercept = 0, slope = -1) +
	geom_vline(xintercept = avgGoals) +
	geom_hline(yintercept = avgGoals) +
	scale_y_reverse()
```

## Change Points to Logos
This requires the ggimage package and changing from geom_point to geom_image
```{r 'Change points to logos'}
library(ggimage)
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.05) +
	geom_abline(intercept = 0, slope = -1) +
	geom_vline(xintercept = avgGoals) +
	geom_hline(yintercept = avgGoals) +
	scale_y_reverse()
```

## Make Axes Same Scale
```{r}
library(ggimage)
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.075) +
	geom_abline(intercept = 0, slope = -1) +
	geom_vline(xintercept = avgGoals) +
	geom_hline(yintercept = avgGoals) +
	coord_fixed(ratio = 1) +
	scale_y_reverse()
```

## Set Title and Axes
There are 3 types of text that need to be set: Plot Title, Axis Titles, and Quadrant Titles. The quadrant titles must be set relative to the visible plot (as opposed to being set based on coordinate location).
```{r}
library(ggimage)
library(grid)
ql.col <- 'red' # Quadrant Label Color
ql.sz <- 10 # Quadrant Label (Font) Size
ql.alpha = 0.3 # Quadrant Label Alpha
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.075) +
	geom_abline(intercept = 0, slope = -1) +
	geom_vline(xintercept = avgGoals) +
	geom_hline(yintercept = avgGoals) +
	coord_fixed(ratio = 1) +
	scale_y_reverse() +
	ggtitle('Team Scoring: Avg Goals For vs Avg Goals Against') +
	xlab('Average Goals For') +
	ylab('Average Goals Against (desc)') +
	annotation_custom(grobTree(textGrob('Good', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Bad', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Exciting', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Boring', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha))))
```

## Set Theme
```{r}
library(ggimage)
library(ggthemes)
ql.col <- 'red' # Quadrant Label Color
ql.sz <- 10 # Quadrant Label (Font) Size
ql.alpha = 0.3 # Quadrant Label Alpha
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.075) +
	geom_abline(intercept = 0, slope = -1) +
	geom_vline(xintercept = avgGoals) +
	geom_hline(yintercept = avgGoals) +
	coord_fixed(ratio = 1) +
	scale_y_reverse() +
	ggtitle('Team Scoring: Avg Goals For vs Avg Goals Against') +
	xlab('Average Goals For') +
	ylab('Average Goals Against (desc)') +
	annotation_custom(grobTree(textGrob('Good', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Bad', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Exciting', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Boring', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	theme_fivethirtyeight(base_size = 10, base_family = 'sans') +
    scale_color_fivethirtyeight() +
	theme(axis.title = element_text())
```

## Adjust Reference Line Styles
Set 1:1 slope line to red and thin. Set mean lines to blue and dashed.
```{r}
library(ggimage)
library(ggthemes)
ql.col <- 'red' # Quadrant Label Color
ql.sz <- 10 # Quadrant Label (Font) Size
ql.alpha = 0.3 # Quadrant Label Alpha
rl.size = 0.25
avgGoalsLabel <- round(avgGoals, digits = 3)
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.075) +
	geom_abline(intercept = 0, slope = -1, color = 'red', size = rl.size) +
	geom_vline(xintercept = avgGoals, color = 'blue', size = rl.size) +
	geom_hline(yintercept = avgGoals, color = 'blue', size = rl.size) +
	coord_fixed(ratio = 1) +
	scale_y_reverse() +
	labs(
		title = 'Team Scoring: Avg Goals For vs Avg Goals Against',
		subtitle = paste0('Average League Goals: ', avgGoalsLabel),
		caption = 'Teams above/to the right of the red line have a positive goal differential.',
		x = 'Average Goals For',
		y = 'Average Goals Against (desc)'
	) +
	annotation_custom(grobTree(textGrob('Good', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Bad', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Exciting', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Boring', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	theme_fivethirtyeight(base_size = 10, base_family = 'sans') +
    scale_color_fivethirtyeight()+
	theme(axis.title = element_text())
```
## Standardize Grid Lines
The axes are for the same value type (both are goals). The grid lines should be at the same interval for both axes.
```{r}
library(ggimage)
library(ggthemes)
ql.col <- 'red' # Quadrant Label Color
ql.sz <- 10 # Quadrant Label (Font) Size
ql.alpha = 0.3 # Quadrant Label Alpha
rl.size = 0.25
avgGoalsLabel <- round(avgGoals, digits = 3)
ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.075) +
	geom_abline(intercept = 0, slope = -1, color = 'red', size = rl.size) +
	geom_vline(xintercept = avgGoals, color = 'blue', size = rl.size) +
	geom_hline(yintercept = avgGoals, color = 'blue', size = rl.size) +
	coord_fixed(ratio = 1) +
	scale_x_continuous(minor_breaks = seq(from = 0, to = 10, by = 0.25)) +
	scale_y_continuous(minor_breaks = seq(from = 0, to = 10, by = 0.25), trans = 'reverse') +
	labs(
		title = 'Team Scoring: Avg Goals For vs Avg Goals Against',
		subtitle = paste0('Average League Goals: ', avgGoalsLabel),
		caption = 'Teams above/to the right of the red line have a positive goal differential.',
		x = 'Average Goals For',
		y = 'Average Goals Against (desc)'
	) +
	annotation_custom(grobTree(textGrob('Good', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Bad', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Exciting', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Boring', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	theme_fivethirtyeight(base_size = 10, base_family = 'sans') +
    scale_color_fivethirtyeight() +
	theme(axis.title = element_text())

```

## Change "Goals" to "Score"
```{r}
library(ggimage)
library(ggthemes)
ql.col <- 'red' # Quadrant Label Color
ql.sz <- 10 # Quadrant Label (Font) Size
ql.alpha = 0.3 # Quadrant Label Alpha
rl.size = 0.25
avgGoalsLabel <- round(avgGoals, digits = 3)
p <- ggplot(goals, aes(AvgGoalsFor, AvgGoalsAgainst)) +
	geom_image(aes(image = Logo), size = 0.075) +
	geom_abline(intercept = 0, slope = -1, color = 'red', size = rl.size) +
	geom_vline(xintercept = avgGoals, color = 'blue', size = rl.size) +
	geom_hline(yintercept = avgGoals, color = 'blue', size = rl.size) +
	coord_fixed(ratio = 1) +
	scale_x_continuous(minor_breaks = seq(from = 0, to = 10, by = 0.25)) +
	scale_y_continuous(minor_breaks = seq(from = 0, to = 10, by = 0.25), trans = 'reverse') +
	labs(
		title = 'Team Scoring: Avg Goals For vs Avg Goals Against',
		subtitle = paste0('Average League Goals: ', avgGoalsLabel),
		caption = 'Teams above/to the right of the red line have a positive goal differential.',
		x = 'Average Goals For',
		y = 'Average Goals Against (desc)'
	) +
	annotation_custom(grobTree(textGrob('Good', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Bad', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Exciting', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	annotation_custom(grobTree(textGrob('Boring', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
	theme_fivethirtyeight(base_size = 10, base_family = 'sans') +
    scale_color_fivethirtyeight() +
	theme(axis.title = element_text())

plotdata <- ggplot_build(p)
```

## Functionalize Graph and Test
```{r}
team_quality_quadrants <- function(homeTeamIds, awayTeamIds, homeScores, awayScores, homeLogoUrls, awayLogoUrls){
	teamGames <- data.frame(Team = c(homeTeamIds, awayTeamIds),
							ScoreFor = c(homeScores, awayScores),
							ScoreAgainst = c(awayScores, homeScores))

	teamLogos <- data.frame(Team = c(homeTeamIds, awayTeamIds), Logo = c(homeLogoUrls, awayLogoUrls)) %>% distinct()

	scores <- teamGames %>%
		group_by(Team) %>%
		summarise(ScoreFor = sum(ScoreFor),
				  ScoreAgainst = sum(ScoreAgainst),
				  Games = n()) %>%
		mutate(AvgScoreFor = ScoreFor / Games,
			   AvgScoreAgainst = ScoreAgainst / Games) %>%
		inner_join(teamLogos, by = 'Team') %>%
		select(Team, AvgScoreFor, AvgScoreAgainst, Games, ScoreFor, ScoreAgainst, Logo) %>%
		arrange(-AvgScoreFor, -AvgScoreAgainst)

	avgScore <- mean(scores$AvgScoreFor)

	maxDecimalPlaces <- 2
	ql.col <- 'red' # Quadrant Label Color
	ql.sz <- 10 # Quadrant Label (Font) Size
	ql.alpha = 0.3 # Quadrant Label Alpha
	rl.size = 0.25

	avgScoreLabel <- format(round(avgScore, digits = maxDecimalPlaces), nsmall = maxDecimalPlaces)

	p <- ggplot(scores, aes(AvgScoreFor, AvgScoreAgainst)) +
		geom_image(aes(image = Logo), size = 0.075) +
		geom_abline(intercept =  0.0, slope = -1, color = ql.col, size = rl.size) +
		geom_vline(xintercept = avgScore, color = 'blue', size = rl.size) +
		geom_hline(yintercept = avgScore, color = 'blue', size = rl.size) +
		coord_fixed(ratio = 1) +
		scale_x_continuous(breaks = seq(from = 0, to = 200, by = 0.25)) +
		scale_y_continuous(breaks = seq(from = 0, to = 200, by = 0.25), trans = 'reverse') +
		labs(
			title = 'Team Scoring: Avg Score For vs Avg Score Against',
			subtitle = paste0('Average League Score: ', avgScoreLabel),
			caption = 'Teams above/to the right of the red line have a positive goal differential.'
		) +
		xlab('Average Score For') +
		ylab('Average Score Against (desc)') +
		annotation_custom(grobTree(textGrob('Good', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		annotation_custom(grobTree(textGrob('Bad', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		annotation_custom(grobTree(textGrob('Exciting', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		annotation_custom(grobTree(textGrob('Boring', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		theme_fivethirtyeight(base_size = 10, base_family = 'sans') +
		scale_color_fivethirtyeight() +
		theme(axis.title = element_text())

	return(p)
}

testViz <- team_quality_quadrants(g2$HomeTeam, g2$AwayTeam, g2$HomeScore, g2$AwayScore, g2$HomeTeamLogoUrl, g2$AwayTeamLogoUrl)
testViz
```

# Team Game Excitement
A high scoring game with a low margin of victory is generally more exciting for fans to watch. The Team Excitement Quadrants will have Average Goals Scored on the x-axis and Average Victory Margin on the y-axis. This will leave the following four quadrants.

- High Total, Low Margin: Electric
- High Total, High Margin: Slaughter
- Low Total, Low Margin: Battle/Grind
- Low Total, High Margin: Surgical/Tactical


## Create Team Game Excitement Function
Using the base functionality produced above, the function for the Team Game Excitement graphic is produced below

```{r}
team_game_excitement_quadrants <- function(homeTeamIds, awayTeamIds, homeScores, awayScores, homeLogoUrls, awayLogoUrls){
	teamGames <- data.frame(Team = c(homeTeamIds, awayTeamIds),
							ScoreFor = c(homeScores, awayScores),
							ScoreAgainst = c(awayScores, homeScores))

	teamLogos <- data.frame(Team = c(homeTeamIds, awayTeamIds), Logo = c(homeLogoUrls, awayLogoUrls)) %>% distinct()

	teamValues <- teamGames %>%
		group_by(Team) %>%
		summarise(ScoreTotal = sum(ScoreFor + ScoreAgainst),
				  AbsScoreDiff = sum(abs(ScoreFor - ScoreAgainst)),
				  Games = n()) %>%
		mutate(AvgScoreTotal = ScoreTotal / Games,
			   AvgScoreDiff = AbsScoreDiff / Games) %>%
		inner_join(teamLogos, by = 'Team') %>%
		select(Team, AvgScoreTotal, AvgScoreDiff, Games, ScoreTotal, AbsScoreDiff, Logo) %>%
		arrange(-AvgScoreTotal, -AvgScoreDiff)

	avgLeagueScoreTotal <- mean(teamValues$AvgScoreTotal)
	avgLeagueScoreDiff <- mean(teamValues$AvgScoreDiff)

	maxDecimalPlaces <- 3
	ql.col <- 'red' # Quadrant Label Color
	ql.sz <- 10 # Quadrant Label (Font) Size
	ql.alpha = 0.3 # Quadrant Label Alpha
	rl.size = 0.25

	avgScoreTotalLabel <- format(round(avgLeagueScoreTotal, digits = maxDecimalPlaces), nsmall = maxDecimalPlaces)
	avgScoreDiffLabel <- format(round(avgLeagueScoreDiff, digits = maxDecimalPlaces), nsmall = maxDecimalPlaces)

	p <- ggplot(teamValues, aes(AvgScoreTotal, AvgScoreDiff)) +
		geom_image(aes(image = Logo), size = 0.075) +
		geom_vline(xintercept = avgLeagueScoreTotal, color = 'blue', size = rl.size) +
		geom_hline(yintercept = avgLeagueScoreDiff, color = 'blue', size = rl.size) +
		coord_fixed(ratio = 1) +
		scale_x_continuous(breaks = seq(from = 0, to = 200, by = 0.25)) +
		scale_y_continuous(breaks = seq(from = 0, to = 200, by = 0.25), trans = 'reverse') +
		labs(
			title = 'Game Excitment: Scoring Frequency vs Game Closeness',
			subtitle = paste('Average Game Total:', avgScoreTotalLabel, '- Average Game Margin:', avgScoreDiffLabel)
		) +
		xlab('Average Game Total') +
		ylab('Average Absolute Game Score Difference (desc)') +
		annotation_custom(grobTree(textGrob('Electric', x = 0.95, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		annotation_custom(grobTree(textGrob('Tactical', x = 0.05, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		annotation_custom(grobTree(textGrob('Slaughter', x = 0.95, y = 0.05, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		annotation_custom(grobTree(textGrob('Grind', x = 0.05, y = 0.95, gp = gpar(col = ql.col, fontsize = ql.sz, alpha = ql.alpha)))) +
		theme_fivethirtyeight(base_size = 10, base_family = 'sans') +
		scale_color_fivethirtyeight() +
		theme(axis.title = element_text())

	return(p)
}

testViz <- team_game_excitement_quadrants(g1$HomeTeam, g1$AwayTeam, g1$HomeScore, g1$AwayScore, g1$HomeTeamLogoUrl, g1$AwayTeamLogoUrl)
testViz <- team_game_excitement_quadrants(g2$HomeTeam, g2$AwayTeam, g2$HomeScore, g2$AwayScore, g2$HomeTeamLogoUrl, g2$AwayTeamLogoUrl)
testViz
```